---
title: "Mediation with R/intermediate"
author: "Brian S. Yandell"
date: "4/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(intermediate)
library(dplyr)
library(ggplot2)
```

```{r}
data(Tmem68)
```

## Mediation scan

```{r}
# target and Tmem68$mediator[,"Tmem68"] are identical.
# so add some noise to make more interesting.
target <- Tmem68$target
target <- target + rnorm(length(target))
```

```{r}
covar <- Tmem68$covar
m <- match("Tmem68", Tmem68$annotation$symbol)
annot_tar <- Tmem68$annotation[m,, drop = FALSE]
```

Reconstruct 8-allele genotype probabilities.

```{r}
driver <- cbind(A = 1 - apply(Tmem68$qtl.geno, 1, sum), Tmem68$qtl.geno)
rownames(driver) <- rownames(Tmem68$qtl.geno)
```

```{r}
annotation <- Tmem68$annotation
mediators <- Tmem68$mediator
```

```{r}
med_scan <- mediation_scan(target = target,
                      mediator = mediators,
                      driver = driver,
                      annotation = annotation,
                      covar = covar,
                      method = "double-lod-diff")
ggplot_mediation_scan(med_scan)
```

Or use `autoplot`, and maybe focus on one group and add a vertical line.

```{r}
autoplot(subset(med_scan, "4")) +
  geom_vline(xintercept = annotation[m,"pos"], linetype = "dashed")

```

## Mediation test

First fine mediators that have significant LOD. This will be used to filter traits to set of potential mediators, and provide annotaton for those mediators.

```{r}
med_lod <- mediator_lod(mediator = mediators,
                        driver = driver,
                        annotation = annotation,
                        covar_med = covar)
med_signif <- med_lod$id[med_lod$lod >= 5]
# Add info column.
med_lod$info <- paste("chr =", med_lod$chr)
```

Causal model selection tests.

```{r}
med_test <- mediation_test(target = target,
                      mediator = mediators[, med_signif, drop = FALSE],
                      driver = driver,
                      annotation = med_lod,
                      covar_tar = covar,
                      covar_med = covar,
                      method = "double-lod-diff")
(sum_med <- summary(med_test) %>%
    arrange(pvalue))
```

```{r}
autoplot(med_test)
```

## Driver effects

```{r}
meds <- 
  dplyr::rename(
    dplyr::mutate(
      dplyr::select(
        sum_med,
        symbol, chr, pos, mediation, pvalue, dplyr::everything()),
      mediation = mediation / log(10),
      chr = as.character(chr),
      LRmed = LRmed / log(10)),
    lod_adj = "mediation",
    mediator = "symbol",
    lod_med = "LRmed")
```

```{r}
driver_levels <- LETTERS[1:8]
```

```{r}
peaks <- fitDefault(driver, target,, covar)
peaks <- data.frame(target = "target",
                    pos_tar = annot_tar$pos,
                    lod_tar = peaks$LR / log(10),
                    group = "1",
                    as.data.frame(t(peaks$coef[driver_levels])),
                    stringsAsFactors = FALSE)
m <- match(driver_levels, names(peaks))
names(peaks)[m] <- paste0(driver_levels, "_p")
```

```{r}
out <- 
  data.frame(meds, peaks) %>%
  dplyr::filter(pvalue <= 0.05)
```

```{r}
plotly::ggplotly(
  ggplot2::ggplot(out) +
    ggplot2::aes(lod_tar - lod_adj, -log10(pvalue), col = triad,
                 symbol = mediator, pvalue = pvalue) +
    ggplot2::geom_point())
```

```{r}
tmp <- rep(seq_len(ceiling(nrow(out) / 6)), each = 6, length = nrow(out))
outp <- split(out, tmp)
driver_effects <- purrr::map(outp, intermediate::driver_effect)
```

```{r fig.width = 12, fig.heigth = 9}
invisible(sapply(driver_effects, function(x) print(intermediate::ggplot_driver_effect(x))))
```

## Mediation triad

```{r}
m <- match("Tmem68", annotation$symbol)
mediator <- mediators[, m, drop = FALSE]
colnames(mediator) <- "Tmem68"
```

```{r}
med_triad <- mediation_triad(target = target,
                      mediator = mediator,
                      driver = driver,
                      covar_tar = covar,
                      sdp = 2)
autoplot(med_triad)
```

